The importance of measurement uncertainty in ecological management

### 2018-01-18 appendix/appendix_harvestfirst.Rmd
WIP

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: medium prior",
                      "biomass",
                      "catch"))
```

@@

```{r}
sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)

@@

```{r}
i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
m <- models[[i]]
alpha <- alphas[[i]] # we need the corresponding alpha vectors
```

@@

```{r}
unif_prior = rep(1, length(states)) / length(states) # initial belief
unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)

@@

star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)

@@

write_csv(all_policies, file.path(log_dir, "all_policies.csv"))

@@

```{r}
pomdp_policies %>% 
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

```{r}
pomdp_policies %>% 
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

```{r}
all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "msy")) %>%
  mutate(policy = fct_recode(policy,
                             "MSY" = "msy",
                             "Constant escapement" = "det",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

```{r}
all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "msy")) %>%
  mutate(policy = fct_recode(policy,
                             "MSY" = "msy",
                             "Constant escapement" = "det",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) %>%
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

>>>>

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))
```

@@

```{r}
sim_col_types <-  "inicnnn"
sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)

@@

```{r}
i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.15) %>% pull(scenario) %>% as.integer()
m <- models[[i]]
alpha <- alphas[[i]] # we need the corresponding alpha vectors  
```

@@

```{r}
unif_prior = rep(1, length(states)) / length(states) # initial belief
unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)
```

@@

star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)
```

@@

write_csv(all_policies, file.path(log_dir, "all_policies.csv"))
```

```{r}
priors <- 
data_frame(states, 
           prior_low,
           prior_star,
           unif_prior,
           prior_high) 
write_csv(priors, file.path(log_dir, "priors.csv"))
```

@@

POMDP policies depend on the prior belief.

```{r fig.width=5, fig.height=7, fig.cap=""}
priors <- 
  read_csv(file.path(log_dir, "priors.csv")) %>%
  select(states, 
         "POMDP: low prior" = prior_low, 
         "POMDP: high prior" = prior_high) %>%
  gather(prior, belief, -states)

policy_df <- all_policies %>% 
  filter(policy %in% c("low", "high", "det", "pgy")) %>%
  mutate(policy = fct_recode(policy,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP: low prior" = "low",
                             "POMDP: high prior" = "high"
                             )) 
prior <- 
  ggplot(priors,aes(states, belief, color = prior)) + 
  geom_line(lwd=1)  + 
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1)) +
  ggtitle("POMDP priors") + 
  xlab("Estimated stock biomass") + 
  theme(legend.position = "none")

harvest <- policy_df %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
  ggtitle("Harvest view") + 
  theme(legend.position="none") + 
  xlab("Estimated stock biomass")

escapement <- policy_df %>%
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
  xlab("Estimated stock biomass") +
  ggtitle("Escapement view") + 
  theme(legend.position = "bottom")

gridExtra::grid.arrange(harvest, escapement, prior,
  layout_matrix = rbind(1,1,2,2, 3))

### 2018-01-18 manuscripts/manuscript.Rmd
WIP

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: medium prior",
                      "biomass",
                      "catch"))

@@

While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach. This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.  The POMDP policy cannot be specified by the most recent observation alone. Instead, the POMDP policy depends on all prior observations, not just the most recent.  The reason for this complexity comes from the Markov property.  Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements. When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.  POMDP formalizes this intutition by capturing the information from all previous observations into a *prior belief*.  This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.  The mechanics of this process are well documented in the extensive literature on POMDPs [e.g. ], but for our purposes

@@

```{r fig.width=7, fig.height=7, fig.cap=""}
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "pgy")) %>%
  mutate(policy = fct_recode(policy,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) -> policy_df
harvest <- policy_df %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
  ggtitle("Harvest view") + 
  theme(legend.position="none") + 
  xlab("Estimated stock biomass")

@@

gridExtra::grid.arrange(harvest, escapement, 
                        layout_matrix = rbind(c(1,1), c(2,2)))

>>>>

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))

@@

Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.  While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach. This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.  The POMDP policy cannot be specified by the most recent observation alone. Instead, the POMDP policy depends on all prior observations, not just the most recent.  The reason for this complexity comes from the Markov property.  Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements. When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.  POMDP formalizes this intutition by capturing the information from all previous observations into a *prior belief*.  This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.  The mechanics of this process are well documented in the extensive literature on POMDPs [e.g. @pomdpstuff; @stuff; @more], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.

Figure 2 shows two separate policy curves for the same POMDP solution given two different prior belief distributions. The lower prior

@@

```{r fig.width=7, fig.height=7, fig.cap=""}
priors <- 
  read_csv(file.path(log_dir, "priors.csv")) %>%
  select(states, 
         "POMDP: low prior" = prior_low, 
         "POMDP: high prior" = prior_high) %>%
  gather(prior, belief, -states)

policy_df <- all_policies %>% 
  filter(policy %in% c("low", "high", "det", "pgy")) %>%
  mutate(policy = fct_recode(policy,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP: low prior" = "low",
                             "POMDP: high prior" = "high"
                             )) 
prior <- 
  ggplot(priors,aes(states, belief, color = prior)) + 
  geom_line(lwd=1)  + 
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1)) +
  ggtitle("POMDP priors") + 
  xlab("Estimated stock biomass") + 
  theme(legend.position = "none")

harvest <- policy_df %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
  ggtitle("Harvest view") + 
  theme(legend.position="none") + 
  xlab("Estimated stock biomass")

@@

gridExtra::grid.arrange(harvest, escapement, prior,
  layout_matrix = rbind(1,1,2,2, 3))
```
